# Conventional combine tests

If we combine certain conventional tests together as a single test, i.e. use the extreme p-value as the final p-value (see [@fisher1992statistical]), we need to adjust the significance level for each individual test.

Considering a combined test consisting of a RESET test and a BP test, we can estimate and plot their $p$-values against each other based on the simulated data.


## Load libraries

```{r}
library(tidyverse)
library(visage)
```

## Load data

### Visual test data

Data with uniform distribution will be used similar to the paper. 

```{r}
visual_test_dat <- vi_survey %>%
  filter(x_dist == "uniform", !attention_check, !null_lineup) %>%
  select(unique_lineup_id, effect_size, type, p_value) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first))
```

```{r}
head(visual_test_dat)
```


### Conventional test data

The simulation data is borrowed from the paper.

```{r}
poly_conv_sim <- readRDS("data/poly_conventional_simulation.rds")
heter_conv_sim <- readRDS("data/heter_conventional_simulation.rds")

# Borrow effect size from the survey
poly_conv_sim <- poly_conv_sim %>%
  left_join(select(filter(vi_survey, type == "polynomial"), 
                   shape, e_sigma, n, x_dist, effect_size))

heter_conv_sim <- heter_conv_sim %>%
  left_join(select(filter(vi_survey, type == "heteroskedasticity"), 
                   a, b, n, x_dist, effect_size))
```


```{r}
conv_test_dat <- bind_rows(poly_conv_sim, heter_conv_sim) %>%
  filter(x_dist == "uniform")
head(conv_test_dat)
```

## Compute GLM for visual test

### Define the minimum and maximum effect size

```{r}
min_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  min()

max_es <- vi_survey %>% 
  filter(!null_lineup, 
         !attention_check,
         x_dist == "uniform") %>%
  pull(effect_size) %>%
  max()
```

```{r}
c(log(min_es), log(max_es))
```



### Fit the model

```{r}
visual_mod <- visual_test_dat %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  mutate(reject = p_value <= 0.05) %>%
  
  # Slope-only model
  glm(reject ~ effect_size - 1,
      family = binomial(),
      data = .,
      offset = offset0)
```

```{r}
summary(visual_mod)
```


### Make prediction for the visual model

```{r}
visual_pred <- data.frame(effect_size = exp(seq(log(min_es), 
                                                log(max_es), 
                                                0.1))) %>%
  mutate(power = predict(visual_mod, 
                         type = "response",
                         newdata = data.frame(effect_size = effect_size,
                                              offset0 = log(0.05/0.95)))) %>%
  mutate(log_effect_size = log(effect_size))
```

```{r}
head(visual_pred)
```

### Make bootstrap prediction for the visual model

```{r cache = TRUE}
visual_boot_pred <- map_dfr(1:500, function(boot_id) {
  
  boot_mod <- visual_test_dat %>%
    mutate(offset0 = log(0.05/0.95)) %>%
    mutate(reject = p_value <= 0.05) %>%
    slice_sample(n = nrow(.), replace = TRUE) %>%
    update(visual_mod, data = .)
  
  data.frame(effect_size = exp(seq(log(min_es), 
                                   log(max_es), 
                                   0.1))) %>%
  mutate(power = predict(boot_mod, 
                         type = "response",
                         newdata = data.frame(effect_size = effect_size,
                                              offset0 = log(0.05/0.95)))) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(boot_id = boot_id)
})

```


## Estimate p-values with simulated data

```{r cache = TRUE}
set.seed(10086)

stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

conv_null_dat <- map_dfr(1:100000, function(i) {
  
  x_dist <- sample(c("uniform", 
                     "normal", 
                     "lognormal", 
                     "even_discrete"), 1)
  x <- switch(x_dist,
              uniform = rand_uniform(-1, 1),
              normal = {
                raw_x <- rand_normal(sigma = 0.3)
                closed_form(~stand_dist(raw_x))
                },
              lognormal = {
                raw_x <- rand_lognormal(sigma = 0.6)
                closed_form(~stand_dist(raw_x/3 - 1))
                },
              even_discrete = rand_uniform_d(k = 5, even = TRUE))
        
  mod <- poly_model(include_z = FALSE, x = x)
  
  n <- sample(c(50, 100, 300), 1)
  
  tmp_dat <- mod$gen(n)
  
  tibble(x_dist = x_dist,
         n = n,
         F_p_value = mod$test(tmp_dat)$p_value,
         RESET_p_value = mod$test(tmp_dat, 
                                  test = "RESET", 
                                  power = 2:4, 
                                  power_type = "fitted")$p_value,
         BP_p_value = HETER_MODEL$test(tmp_dat)$p_value,
         SW_p_value = shapiro.test(tmp_dat$.resid)$p.value,
         boot_id = i)
})
```


```{r}
conv_null_dat %>%
  ggplot() +
  geom_density2d_filled(aes(RESET_p_value, BP_p_value)) +
  scale_x_sqrt() +
  scale_x_sqrt() +
  xlab("RESET p-value") +
  ylab("BP p-value")
```


We can pick a rectangle area at the bottom left corner such that $P(\text{RESET rejects } H_0 \text{ or BP rejects } H_0) = \alpha$. The side length of this rectangle is denoted as $\alpha^*$.

```{r}
alpha_star <- map_dbl(seq(0.0001, 0.05, 0.0005), function(alpha_star) {
  conv_null_dat %>%
    filter(RESET_p_value <= alpha_star | BP_p_value <= alpha_star) %>%
    nrow()/nrow(conv_null_dat)
}) %>%
  {seq(0.0001, 0.05, 0.0005)[which.min(abs(. - 0.05))]}
```

```{r}
alpha_star
```

## Fit GLM

```{r}
conv_test_group_dat <- conv_test_dat %>%
  mutate(`Group test (RESET & BP)_p_value` = ifelse(RESET4_p_value <= alpha_star | BP_p_value <= alpha_star, 0, 1)) %>%
  relocate(`Group test (RESET & BP)_p_value`, .after = SW_p_value)
```

```{r cache = TRUE}
conv_pred_group <- conv_test_group_dat %>%
  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  pivot_longer(RESET_p_value:`Group test (RESET & BP)_p_value`) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    data.frame(effect_size = exp(seq(log(min_es),
                                     log(max_es),
                                     0.1)),
               offset0 = log(0.05/0.95)) %>%
      mutate(power = predict(mod, type = "response", newdata = .))
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  select(-offset0) %>%
  mutate(log_effect_size = log(effect_size))
```

## Draw plot

```{r}
ggplot() +
  geom_point(data = visual_test_dat,
             aes(log(effect_size), as.numeric(p_value <= 0.05)),
             alpha = 0.15) +
  
  geom_line(data = conv_pred_group,
            aes(log_effect_size, power, col = name),
            size = 1) +
  
  geom_line(data = visual_boot_pred,
            aes(log_effect_size, power, col = "visual test", group = boot_id),
            size = 1,
            alpha = 0.01) +
  
  geom_line(data = visual_pred,
            aes(log_effect_size, power, col = "visual test"), 
            size = 1) +
  
  theme_light() +
  scale_color_manual(values = rev(rcartocolor::carto_pal(5, "Vivid"))) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

The combined test has higher power than the visual test in this case.

# Combined tests (RESET, BP and SW)

Let's consider different combined tests consisting of the RESET test, BP test and the SW test.

```{r}
alpha_star2 <- map_dbl(seq(0.0001, 0.05, 0.0005), function(alpha_star) {
  conv_null_dat %>%
    filter(RESET_p_value <= alpha_star | SW_p_value <= alpha_star) %>%
    nrow()/nrow(conv_null_dat)
}) %>%
  {seq(0.0001, 0.05, 0.0005)[which.min(abs(. - 0.05))]}
```

```{r}
alpha_star2
```


```{r}
alpha_star3 <- map_dbl(seq(0.0001, 0.05, 0.0005), function(alpha_star) {
  conv_null_dat %>%
    filter(BP_p_value <= alpha_star | SW_p_value <= alpha_star) %>%
    nrow()/nrow(conv_null_dat)
}) %>%
  {seq(0.0001, 0.05, 0.0005)[which.min(abs(. - 0.05))]}
```

```{r}
alpha_star3
```

```{r}
alpha_star4 <- map_dbl(seq(0.0001, 0.05, 0.0005), function(alpha_star) {
  conv_null_dat %>%
    filter(RESET_p_value <= alpha_star | BP_p_value <= alpha_star | SW_p_value <= alpha_star) %>%
    nrow()/nrow(conv_null_dat)
}) %>%
  {seq(0.0001, 0.05, 0.0005)[which.min(abs(. - 0.05))]}
```

```{r}
alpha_star4
```

## Fit GLM

```{r}
conv_test_group_dat <- conv_test_dat %>%
  mutate(`Group (RESET & BP)_p_value` = ifelse(RESET4_p_value <= alpha_star | BP_p_value <= alpha_star, 0, 1)) %>%
  relocate(`Group (RESET & BP)_p_value`, .after = SW_p_value) %>%
  
  mutate(`Group (RESET & SW)_p_value` = ifelse(RESET4_p_value <= alpha_star2 | SW_p_value <= alpha_star2, 0, 1)) %>%
  relocate(`Group (RESET & SW)_p_value`, .after = `Group (RESET & BP)_p_value`) %>%
  
  mutate(`Group (BP & SW)_p_value` = ifelse(BP_p_value <= alpha_star3 | SW_p_value <= alpha_star3, 0, 1)) %>%
  relocate(`Group (BP & SW)_p_value`, .after = `Group (RESET & SW)_p_value`) %>%
  
  mutate(`Group (RESET & BP & SW)_p_value` = ifelse(RESET4_p_value <= alpha_star3 | BP_p_value <= alpha_star3 | SW_p_value <= alpha_star3, 0, 1)) %>%
  relocate(`Group (RESET & BP & SW)_p_value`, .after = `Group (BP & SW)_p_value`)
```


```{r cache = TRUE}
conv_pred_group <- conv_test_group_dat %>%
  select(-RESET3_p_value, -(RESET5_p_value:RESET10_p_value), -F_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  pivot_longer(RESET_p_value:`Group (RESET & BP & SW)_p_value`) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value <= 0.05) %>%
  select(effect_size, name, reject) %>%
  mutate(offset0 = log(0.05/0.95)) %>%
  nest(dat = c(effect_size, offset0, reject)) %>%
  mutate(mod = map(dat, 
                   ~glm(reject ~ effect_size - 1, 
                        family = binomial(), 
                        data = .x,
                        offset = offset0))) %>%
  mutate(power = map(mod, function(mod) {
    data.frame(effect_size = exp(seq(log(min_es),
                                     log(max_es),
                                     0.1)),
               offset0 = log(0.05/0.95)) %>%
      mutate(power = predict(mod, type = "response", newdata = .))
  })) %>%
  select(-dat, -mod) %>%
  unnest(power) %>%
  select(-offset0) %>%
  mutate(log_effect_size = log(effect_size))
```

```{r}
conv_pred_group <- conv_pred_group %>%
  mutate(name = factor(name, levels = c("Group (RESET & BP & SW) test",
                                        "Group (RESET & BP) test",
                                        "Visual test",
                                        "Group (RESET & SW) test",
                                        "Group (BP & SW) test",
                                        "BP test",
                                        "SW test",
                                        "RESET test")))
```


## Draw plot

```{r}
ggplot() +
  geom_point(data = visual_test_dat,
             aes(log(effect_size), as.numeric(p_value <= 0.05)),
             alpha = 0.15) +
  
  geom_line(data = conv_pred_group,
            aes(log_effect_size, power, col = name),
            size = 1) +
  
  geom_line(data = visual_boot_pred,
            aes(log_effect_size, power, col = "Visual test", group = boot_id),
            size = 1,
            alpha = 0.01) +

  geom_line(data = visual_pred,
            aes(log_effect_size, power, col = "Visual test"),
            size = 1) +
  
  theme_light() +
  # scale_color_brewer(palette = "Set3") +
  scale_color_manual(breaks = levels(conv_pred_group$name), 
                     values = rev(rcartocolor::carto_pal(8, "Vivid"))) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups") +
  theme(legend.position = "bottom")
```


The sensitivity of the tests are ordered as follows: 

- Group test (RESET & BP & SW)
- Group test (RESET & BP)
- Visual test
- Group test (RESET & SW) test
- Group test (BP & SW) test
- BP test
- SW test
- RESET test


